### 2018-04-10 latex/icpp18/abstract.tex
submitted fist draft icpp18

\begin{abstract}
Non-negative matrix factorization (NMF), the problem of finding two non-negative low-rank factors whose product approximates an input matrix, is a useful tool for many data mining and scientific applications such as topic modeling in text mining and blind source separation in microscopy.
In this paper, we focus on scaling algorithms for NMF to very large sparse datasets and massively parallel machines by employing effective algorithms, communication patterns, and partitioning schemes that leverage the sparsity of the input matrix.
In particular, we consider two previous works developed for related problems, one that uses a fine-grained partitioning strategy using a point-to-point communication pattern and on that uses a checkerboard partitioning strategy using a collective-based communication pattern.
We show that a combination of the previous approaches balances the demands of the various computations within NMF algorithms and achieves high efficiency and scalability.
\grey{To be completed...

We present experiments on (THESE DATASETS), running on up to (YY) cores of a Cray XK7 supercomputer.
(MAYBE STATE SIMPLE VERSION OF CONCLUSIONS?)
}

>>>>

\begin{abstract}
Non-negative matrix factorization (NMF), the problem of finding two non-negative low-rank factors whose product approximates an input matrix, is a useful tool for many data mining and scientific applications such as topic modeling in text mining and blind source separation in microscopy.
In this paper, we focus on scaling algorithms for NMF to very large sparse datasets and massively parallel machines by employing effective algorithms, communication patterns, and partitioning schemes that leverage the sparsity of the input matrix.
In the case of machine learning workflow, the 
computations after SpMM must deal with dense matrices, as Sparse-Dense matrix multiplication will result in a dense matrix. 
Hence, the partitioning strategy considering only SpMM will result in a huge imbalance in the overall workflow especially on computations after SpMM and in this specific case of NMF on non-negative least squares computations. 
Towards this, we consider two previous works developed for related problems, one that uses a fine-grained partitioning strategy using a point-to-point communication pattern and on that uses a checkerboard partitioning strategy using a collective-based communication pattern.
We show that a combination of the previous approaches balances the demands of the various computations within NMF algorithms and achieves high efficiency and scalability. From the experiments, we could see that our proposed algorithm communicates atleast
4x less than the collective and achieves upto 100x speed over the baseline FAUN on real world datasets. Our algorithm was experimented in two different super computing platforms and we could scale up to 32000 processors on Bluegene/Q.

### 2018-04-10 latex/icpp18/experiments.tex
submitted fist draft icpp18

\subsection {Effect of communication scheme}

\subsection {Effect of partitioning}

@@

\caption{Communication and computation costs of different partitioning strategies on Kronecker graphs}
\end{figure}

>>>>

%\subsection {Effect of communication scheme}

\subsection {Effect of partitioning}
To understand the effect of partitioning and its impact on communication and computation, we performed
experiments on synthetic Kronecker graph and the results are presented in Figure \ref{fig:kronecker}. 
For $2^{20}$ vertices, we chose the edge factor of the kronecker graph as \{4,8,16,32,64\} and in
this setting as the edge factor increases,  higher the number of  non-zeros making the sparse matrix
denser. We ran the baseline FAUN algorithm and the proposed algorithm with different partitioning 
scheme as explained in Section \ref{sec:partitioning} on 4096 processors in Rhea for low rank $k=48$. 
From the Figure \ref{fig:nnzsweepcomm},  it is evident that checkerboard partitioning with point-to-point
communication is approximately 4x faster over the 2D partition with collective communication. In the case of
computation, from Figure \ref{fig:nnzsweepcomp}, we can observe that in the case of denser matrices
with higher edge factors, the load imbalance in 2D is becoming more evident over the checkerboard partition. 
Thus checker board with point-to-point communicates less and has better load balance over 2D partition
with collective communication. These advantages  give scope for the checkerboard partition with point-to-point scale
to higher number of nodes.

@@

\caption{Communication and computation costs of different partitioning strategies on Kronecker graphs on 4096 processors Rhea for $k=48$}
\label{fig:kronecker}
\end{figure}
